% \documentclass[sigconf, review, anonymous]{acmart}
% \documentclass[sigconf, review]{acmart}
\documentclass[sigconf]{acmart}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

\usepackage{algorithmic}
\usepackage[ruled,linesnumbered]{algorithm2e}
\usepackage{array}
\usepackage{amsfonts}
\let\Bbbk\relax\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{color}
\usepackage{epstopdf}
\usepackage{enumitem}
\usepackage{footmisc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{multirow}
\usepackage{mathrsfs}
\usepackage{pifont}
\usepackage{subfigure}
\usepackage{setspace}
\usepackage{soul}
\usepackage{textcomp}
\usepackage{url}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage[np, autolanguage]{numprint}
\usepackage[skip=0pt]{caption}
\usepackage{marvosym}
\let\Cross\relax
\usepackage{bbding}

\newcommand{\paratitle}[1]{\vspace{1.5ex}\noindent\textbf{#1}}
\newcommand{\ie}{\emph{i.e.,}\xspace}
\newcommand{\aka}{\emph{a.k.a.,}\xspace}
\newcommand{\eg}{\emph{e.g.,}\xspace}
\newcommand{\etal}{\emph{et al.}\xspace}
\newcommand{\wrt}{w.r.t.\xspace}
\newcommand{\ignore}[1]{}
\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\changed}[1]{\textcolor{blue}{#1}}
\newcommand{\remove}[1]{\sout{\textcolor{orange}{#1}}}
\newcommand{\replace}[2]{\remove{#1}/\changed{#2}}
\newcommand\mathbox[1]{\mbox{$#1$}}
\newcommand{\tabincell}[2]{\begin{tabular}{@{}#1@{}}#2\end{tabular}}
\newcommand{\placeholder}[1]{\textcolor{blue}{#1}}
\newcommand{\shanlei}[1]{\textcolor{purple}{#1}}
\newcommand{\modelname}{MDL}

\begin{document}

\title{MDL: A Unified Multi-Distribution Learner in Large-scale Industrial Recommendation through Tokenization}


\author{Shanlei Mu}
\email{mushanlei@bytedance.com}
\authornote{Equal contribution.}
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}

\author{Yuchen Jiang}
\email{jiangyuchen.jyc@bytedance.com}
\authornotemark[1]
\affiliation{
    \institution{ByteDance AML}
    \city{Beijing}
    \country{China}
}

\author{Shikang Wu}
\email{wushikang@bytedance.com}
\authornotemark[1]
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}

\author{Shiyong Hong}
\email{hongshiyong.66@bytedance.com}
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}


\author{Tianmu Sha}
\email{shatianmu@bytedance.com}
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}

\author{Junjie Zhang}
\email{zhangjunjie.leo@bytedance.com}
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}

\author{Jie Zhu}
\email{zhujie.zj@bytedance.com}
\affiliation{
    \institution{ByteDance AML}
    \city{Hangzhou}
    \country{China}
}

\author{Zhe Chen}
\email{chenzhe.john@bytedance.com}
\affiliation{
    \institution{ByteDance AML}
    \city{Beijing}
    \country{China}
}

\author{Zhe Wang}
\authornote{Corresponding author.}
\email{zhewang.tim@gmail.com}
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}

\author{Jingjian Lin}
\email{linjingjian000@gmail.com}
\affiliation{
    \institution{ByteDance Search}
    \city{Beijing}
    \country{China}
}

\renewcommand{\shortauthors}{Shanlei Mu, et al.}

\begin{abstract}
Industrial recommender systems increasingly adopt multi-scenario learning (MSL) and multi-task learning (MTL) to handle diverse user interactions and contexts, but existing approaches suffer from two critical drawbacks: (1) underutilization of large-scale model parameters due to limited interaction with complex feature modules, and (2) difficulty in jointly modeling scenario and task information in a unified framework.
To address these challenges, we propose a unified \textbf{M}ulti-\textbf{D}istribution \textbf{L}earning (MDL) framework, 
inspired by the "prompting" paradigm in large language models (LLMs).
MDL treats scenario and task information as specialized tokens rather than auxiliary inputs or gating signals.
Specifically, we introduce a unified information tokenization module that transforms features, scenarios, and tasks into a unified tokenized format.
To facilitate deep interaction, we design three synergistic mechanisms: (1) feature token self-attention for rich feature interactions, (2) domain-feature attention for scenario/task-adaptive feature activation, and (3) domain-fused aggregation for joint distribution prediction.
By stacking these interactions, MDL enables scenario and task information to "prompt" and activate the model's vast parameter space in a bottom-up, layer-wise manner.
Extensive experiments on real-world industrial datasets demonstrate that MDL significantly outperforms state-of-the-art MSL and MTL baselines.
Online A/B testing on Douyin Search platform over one month yields +0.0626\% improvement in LT30 and -0.3267\% reduction in change query rate.
MDL has been fully deployed in production, serving hundreds of millions of users daily.
\end{abstract}

\begin{CCSXML}
<ccs2012>
 <concept>
 <concept_id>10002951.10003317.10003347.10003350</concept_id>
 <concept_desc>Information systems~Recommender systems</concept_desc>
 <concept_significance>500</concept_significance>
 </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Recommender systems}

\keywords{Recommender System; Large Recommendation Model; Multi-Scenario Learning; Multi-Task Learning; Tokenization}


\maketitle

\input{subfile/sec-intro}
\input{subfile/sec-pre}
\input{subfile/sec-model}
\input{subfile/sec-exp}
\input{subfile/sec-rel}

\section{Conclusion}
\label{sec-con}
In this paper, we propose a novel multi-distribution learning framework for multi-scenario learning (MSL) and multi-task learning (MTL) in large-scale industrial recommender system, named \textbf{MDL}.
Unlike existing MSL and MTL recommendation methods that are based on a \emph{shared-specific} architecture, the proposed approach inspired by large language models (LLMs) is to leverage prompt-based tokenization to achieve model predictions in various scenarios and tasks settings.
Specifically, MDL tokenizes feature, scenario, and task information into a unified tokenized format, and elobrates three key interaction mechanisms among these tokens.
Extensive experiments on both offline evaluation and online test have demonstrated the effectiveness.
The proposed approach MDL has been fully deployed on Douyin Search platform.

\balance
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

\end{document}
